Pandemic Pass? Treaty Derogations and Human Rights Practices During COVID-19

### 2023-03-03 manuscript/manuscript.qmd
super minor changes

For our second hypothesis, we use a measure of civil society repression from the Varieties of Democracy (V-Dem) project [@vdem-v12], derived from a latent Bayesian item response theory model of expert opinions, with values ranging from roughly −3 to 3 (where higher values represent less repression). This data is only recorded annually, so as with our PanDem-based outcomes, civil society repression reflects the overall anti=civil society environment.

@@

In contrast, the patterns in predicted probabilities of the adherence to time limit measures are influenced strongly by derogation status. For non-derogating countries, the two most likely predicted outcomes throughout the beginning of the pandemic are to have no issues with time limits (with a roughly 60% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with 30–40% probability). The middle ground of only minor issues with time limits is exceptionally unlikely. Countries that formally derogated, on the other hand, have a 90% probability of having no issues with time limits throughout the 15 months, and only a less-than-10% probability of moderate issues. The derogation process requires that emergency measures be timebound and limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.

>>>>

For our second hypothesis, we use a measure of civil society repression from the Varieties of Democracy (V-Dem) project [@vdem-v12], derived from a latent Bayesian item response theory model of expert opinions, with values ranging from roughly −3 to 3 (where higher values represent less repression). This data is only recorded annually, so as with our PanDem-based outcomes, civil society repression reflects the overall anti-civil society environment.

@@

In contrast, the patterns in predicted probabilities of the adherence to time limit measures are influenced strongly by derogation status. For non-derogating countries, the two most likely predicted outcomes throughout the beginning of the pandemic are to have no issues with time limits (with a roughly 60% probability) and to have moderate issues with limits, or an absence of an end date for emergency measures (with 30–40% probability). The middle ground of only minor issues with time limits is exceptionally unlikely. Countries that formally derogated, on the other hand, have a 90% probability of having no issues with time limits throughout the 15 months, and only a less-than-10% probability of moderate issues. The derogation process requires that emergency measures be time-limited, and countries that care about their treaty obligations appear to actually follow the time limits they establish.
